Bricking up and overgrowing backdoors in Intrusion Detection Systems

### 2019-09-09 ex2/paper/paper.tex
Modified paper

\section{Introduction}
The detection of attacks in data networks is a fundamental task in data security. Due to the considerable amount of data which has to be analyzed the use of machine learning techniques for this purpose seems natural and is increasingly deployed.

For research, the invention and assessment of techniques for network anomaly detection poses many challenges. To implement such a system, a considerable number of features can be extracted from network data which might be beneficial for anomaly detection and it is not always clear which features are actually important.

For the training of models usually datasets are used which have been generated artificially in a controlled test environment. As a downside of this approach, it is unclear whether a machine learning model learns to classify based on characteristics that are inherent to the attacks which should be detected, or rather learns to classify based on patterns that were unintentionally created during dataset generation.

@@

% TODO: think meghdouri_analysis_2018 didnt use this
In this document, we redo parts of a recent paper~\cite{meghdouri_analysis_2018} which bases on the CIC-IDS-2017 dataset \cite{sharafaldin_toward_2018} for evaluating the performance of several feature vectors and machine learning techniques for accurate anomaly detection. We use explainability methods for investigating if the decisions the anomaly detectors untertake are reasonable.

Furthermore, we add a backdoor to the trained model and show that attack detection can efficiently be bypassed if the attacker had the ability to modify training data. Finally, we apply the same explainability methods to the backdoored model and show how such attack attempts might be recognized before any harm is done.

@@

\section{Experimental Setup} \label{sec:ml_approaches}
\subsection{The Backdoor}
Creating a model for anomaly detection in network traffic is a challenging task which involves massive datasets and significant amounts of computation power. In a practical deployment, it is therefore reasonable to assume that the training is done by a security company marketing either a complete anomaly detection system or just a model usable by an anomaly detection software which might even be open-source.

If we have to question if such a security company can be trusted under all circumstances, the problem occurs that the security company might have tried to implement measures to circumvent reliable attack detection of their attack detection system.

For this reason, we now add a backdoor to our deep learning and random forest models and investigate if detection of malicious training can be detected by applying explainability methods to the training models.

@@

However,

>>>>

\section{Introduction}
The detection of attacks in data networks is a fundamental task in data security. Due to the considerable amount of data which has to be analyzed, the use of machine learning techniques for this purpose seems natural and is increasingly deployed.

For research, the invention and assessment of techniques for \glspl{ids} poses many challenges. To implement such a system, a considerable number of features can be extracted from network data which might be beneficial for anomaly detection and it is not always clear which features are actually important.

Creating a model for an \gls{ids} is a challenging task which involves massive datasets and significant amounts of computational power. In a practical deployment, it is therefore reasonable to assume that the training is done by a security company marketing either a complete anomaly detection system or just a model usable by an anomaly detection software which might even be open-source.

If we have to question if such a security company can be trusted under all circumstances, the problem arises that the security company might have tried to implement measures to circumvent reliable attack detection of their attack detection system.

Furthermore, for the training of models usually datasets are used which have been generated artificially in a controlled test environment. As a downside of this approach, it is unclear whether a machine learning model learns to classify based on characteristics that are inherent to the attacks which should be detected, or rather learns to classify based on patterns that were unintentionally created during dataset generation.

@@

% TODO: think meghdouri_analysis_2018 didnt use this
In this paper, we redo parts of a recent paper~\cite{meghdouri_analysis_2018} which bases on the UNSW-NB15 dataset \cite{moustafa_unsw-nb15:_2015} for evaluating the performance of several feature vectors and machine learning techniques for accurate anomaly detection in the context of \glspl{ids}.
%We use explainability methods for investigating if the decisions the anomaly detectors untertake are reasonable.
Furthermore, we add a backdoor to the trained model and show that attack detection can efficiently be bypassed if the attacker had the ability to modify training data.

\newpage
We then discuss several techniques to detect or remove a backdoor from a trained model. In particular, we show how visualization techniques from explainable \gls{ml} can be used to detect backdoors and problems emerging from the distribution of attack samples in the training dataset.

Probably the most important \gls{ml} method in the context of \glspl{ids} are \gls{rf} classifiers. We propose a new pruning technique specifically for removing backdoors from trained \gls{rf} models.

Finally, we also evaluate techniques proposed for \glspl{cnn} for removing backdoors from image classifiers, which, however, surprisingly show to be uneffective for our \gls{mlp} classifiers.

@@

\section{Experimental Setup} \label{sec:ml_approaches}
\subsection{The Backdoor}

We add a backdoor to our deep learning and random forest models and investigate if detection of malicious training can be detected by applying explainability methods to the training models.

@@

Most importantly, however, features to use for deploying an \gls{ids} have to be selected carefully.
